---
title: "CysC Data Cleaning"
author: "César Higgins"
date: "9/8/2021"
output:
  pdf_document: default
  code_folding: hide
  highlight: pygments
  keep_md: no
  number_sections: yes
  smart: yes
  theme: sandstone
  toc: yes
  toc_float: yes
  html_document: null
---

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(compareGroups)
library(knitr)
library(kableExtra)
library(janitor)
library(broom)
library(lubridate)
library(magrittr) # need to run every time you start R and want to use %>%
library(kableExtra)
library(gtsummary)
library(foreign)
library(gmodels)
library(stringr)
library(dplyr)  

summarize = dplyr::summarize

knitr::opts_chunk$set(
  eval       = TRUE,    # whether to run code in code chunk
  include    = TRUE,    # whether to include the chunk output
  echo       = TRUE,   # Whether to show code chunk in final output
  error      = TRUE,    # whether to display error messages
  message    = FALSE,   # whether to preserve messages
  warning    = FALSE,   # whether to preserve warnings
  comment    = "#>",    # a character string to append at start
                        # of each line of results in final document
  tidy       = FALSE,   # whether to tidy code chunks for display
  #dpi        = 96, 
  #fig.width  = 6,       # consistent width for figures
  #fig.asp    = 0.618,   # the golden ratio, can be adjusted in individual chunks
  out.width  = "100%",   # controls the output size
  fig.align  = "center" # give plot room to breathe
)

```



```{r read data, results="hide"}

rm(list = ls())
setwd("~/Documents/Dementia project /NIH grant/Data")
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 
paste0(source_dir, "hrs_democoggen_wide20200910.Rda")
load("hrs_democoggen_wide20200910.Rda") #This was the data set previously used by it was missing one observation :/

dem <- hrs0910

```


# Cleaning Biomarkers Data Sets

```{r read data from DBS}

setwd("~/Documents/Dementia project /NIH grant/Data")
load("biomk06.RData")
load("biomk08.RData")
biomk06 <- biomk06
biomk08 <- biomk08

myvars_bio_2006 <- c("HHID", "PN", "KA1C_ADJ", "KHDL_ADJ", "KTC_ADJ", "KCYSC_ADJ", "KCRP_ADJ") 
myvars_bio_2008 <- c("HHID", "PN", "LA1C_ADJ", "LHDL_ADJ", "LTC_ADJ", "LCYSC_ADJ", "LCRP_ADJ") 

biomk06_f <- dplyr::select(biomk06,myvars_bio_2006)  
biomk08_f <- dplyr::select(biomk08,myvars_bio_2008)

# Renaming Variables for each data set biomarker data 
#2006 
biomk06_f <- dplyr::rename(biomk06_f, a1c_2006=KA1C_ADJ , hdl_2006 = KHDL_ADJ,  tc_2006 = KTC_ADJ, cysc_2006 = KCYSC_ADJ, crp_2006 = KCRP_ADJ)

# 2008 
biomk08_f <- dplyr::rename(biomk08_f, a1c_2008 = LA1C_ADJ, hdl_2008 = LHDL_ADJ, tc_2008 = LTC_ADJ,cysc_2008 = LCYSC_ADJ, crp_2008 = LCRP_ADJ)

```

# Mergin Biomarkers Data Set with Cognition Data Set 

```{r}

merged_data <- left_join(dem, biomk06_f, by = c("HHID" = "HHID", "PN" = "PN")) 
merged_data_f <- left_join(merged_data, biomk08_f, by = c("HHID" = "HHID", "PN" = "PN"))

# Checking the dimensions of the data set to see that everything is fine 


dem %>% 
dim()

biomk06_f %>% 
dim()

biomk08_f %>% 
dim()

merged_data %>% 
dim()

merged_data_f %>% 
dim()
```

# Reducing the Dimension of the Data 

```{r}

myvars <- c("HHID", "PN", "a1c_2006", "hdl_2006", "tc_2006", "cysc_2006", "crp_2006", "a1c_2008", "hdl_2008", "tc_2008", "cysc_2008", "crp_2008", "DEGREE", "GENDER", "SCHLYRS", "HISPANIC", "RACE", "AD95", "AD96", "AD98", "AD00", "AD02", "AD04", "AD06", "AD08", "AD10", "AD12", "AD14", "expert_dem_98", "expert_dem_00", "expert_dem_02", "expert_dem_04", "expert_dem_06", "expert_dem_08", "expert_dem_10", "expert_dem_12", "expert_dem_14", "AAGE", "BAGE", "CAGE", "DAGE", "EAGE", "FAGE", "GAGE", "HAGE", "JAGE", "KAGE", "LAGE", "MAGE", "NAGE", "OAGE", "PAGE","totscore95", "totscore96", "totscore98", "totscore00", "totscore02", "totscore04", "totscore06", "totscore08", "totscore10", "totscore12", "totscore14", "R3HIBPE", "R4HIBPE", "R5HIBPE", "R6HIBPE", "R7HIBPE", "R8HIBPE", "R9HIBPE", "R10HIBPE", "R11HIBPE", "R12HIBPE", "smoke96", "smoke98", "smoke00", "smoke02", "smoke04", "smoke06", "smoke08", "smoke10", "smoke12", "smoke14","R3DIABE", "R4DIABE", "R5DIABE", "R6DIABE", "R7DIABE", "R8DIABE", "R9DIABE", "R10DIABE", "R11DIABE", "R12DIABE", "R3BMI", "R4BMI", "R5BMI", "R6BMI", "R7BMI", "R8BMI", "R9BMI", "R10BMI", "R11BMI", "R12BMI", "APOE2010_bin", "EA_APOE4", "AA_APOE4", "BIRTHYR", "STROKE95",  "STROKE96", "STROKE98", "STROKE00", "STROKE02", "STROKE04", "STROKE06", "STROKE08", "STROKE10", "STROKE12", "STROKE14", "DA009", "EA009", "FA009", "GA009", "HA009", "JA009", "KA009", "LA009", "MA009", "NA009", "OA009", "R8DRINKN", "R9DRINKN")

dem_final <- dplyr::select(merged_data_f,myvars)

dem_final %>% 
  dim()

```

# Creating the Outcome Variable

```{r}
# Data cleaning 

# Assigning Waves & Creating Outcome Variables 
dem_final_f <- dem_final %>% 
  mutate(wave_2006 = case_when(
    (!is.na(a1c_2006) | !is.na(hdl_2006) | !is.na(tc_2006) | !is.na(cysc_2006) | !is.na(crp_2006)) ~ 1, 
    TRUE ~ 0
  )) %>% 
  mutate(wave_2008 = case_when(
    (!is.na(a1c_2008) | !is.na(hdl_2008) | !is.na(tc_2008) | !is.na(cysc_2008) | !is.na(crp_2008)) ~ 1, 
    TRUE ~ 0
  )) %>% 
  mutate(wave = case_when(
    wave_2006==1 ~ 2006,
    wave_2008==1 ~ 2008,
  )) %>% 
  mutate(AD_06 = case_when(
    AD06==3 ~ 0, 
    AD06==2 ~ 1,
    AD06==1 ~ 2
  )) %>% 
  mutate(AD_08 = case_when(
    AD08==3 ~ 0,
    AD08==2 ~ 1, 
    AD08==1 ~ 2
  )) %>% 
  mutate(demdx= case_when(
    wave==2006 ~ AD_06,
    wave==2008 ~ AD_08)) 


```

# Covariates for Analysis 

```{r}

dem_final_ff <-  dem_final_f %>% 
  mutate(age = wave-BIRTHYR) %>% 
  mutate(education = SCHLYRS) %>% 
  mutate(stroke = case_when(
    wave==2006 ~ STROKE06,
    wave==2008 ~ STROKE08
  )) %>% 
  mutate(hbp = case_when(
    wave==2006 ~ R8HIBPE,
    wave==2008 ~ R9HIBPE # had a mistake in here as R8HIBPE, just corrected it 
  )) %>% 
  mutate(t2db = case_when(
    wave==2006 ~ R8DIABE,
    wave==2008 ~ R9DIABE
  )) %>% 
  mutate(bmi = case_when(
    wave==2006 ~ R8BMI,
    wave==2008 ~ R9BMI
  )) %>% 
  mutate(smoking = case_when(
    wave==2006 ~ smoke06,
    wave==2008 ~ smoke08
  )) %>% 
  mutate(alcohol = case_when(
    wave==2006 ~ R8DRINKN,
    wave==2008 ~ R9DRINKN
  )) %>% 
  mutate(proxy = case_when(
    wave==2006 ~ KA009,
    wave==2008 ~ LA009,
  )) %>% 
  mutate(proxy_f= case_when(
    proxy==1 ~ 1,
    proxy==2 ~ 2, 
    proxy==3 ~ 2
  )) %>% 
  mutate(totalscore= case_when(
    wave==2006 ~ totscore06,
    wave==2008 ~ totscore08
  )) %>% 
  mutate(race_f = case_when(
    RACE==0 ~ "NA",
    RACE==1 ~ "White",
    RACE==2 ~ "Black",
    RACE==7 ~ "Other")) %>% 
  mutate(hispanic_f = case_when(
      HISPANIC==1 | HISPANIC==2 | HISPANIC==3 ~ "Hispanic",
      HISPANIC==5 ~ "Non-Hispanic",
      HISPANIC==0 ~ "NA")) %>% 
  mutate(baseline_crp = case_when(
    wave==2006 ~ crp_2006,
    wave==2008 ~ crp_2008
  )) %>% 
  mutate(baseline_hdl= case_when(
    wave==2006 ~ hdl_2006,
    wave==2008 ~ hdl_2008
  )) %>% 
  mutate(baseline_cysc = case_when(
    wave==2006 ~ cysc_2006,
    wave==2008 ~ cysc_2008
  )) %>% 
  mutate(baseline_tc = case_when(
    wave==2006 ~ tc_2006,
    wave==2008 ~ tc_2008
  )) %>% 
  mutate(baseline_a1c = case_when(
    wave==2006 ~ a1c_2006,
    wave==2008 ~ a1c_2008
  ))

 dem_final_fff <- dem_final_ff %>% 
  mutate(race_f = na_if(race_f, "NA")) %>% 
  mutate(hispanic_f = na_if(hispanic_f, "NA")) %>% 
  mutate(race_ff = case_when(
    race_f=="White" & hispanic_f=="Non-Hispanic" ~ "White-NH",
    race_f=="Black" & hispanic_f=="Non-Hispanic" ~ "Black-NH",
    race_f=="Other" & hispanic_f=="Non-Hispanic" ~ "Other-NH",
    hispanic_f=="Hispanic" ~ "Hispanic")) %>% 
  mutate(gender_f = case_when(
    GENDER==1 ~ "Male",
    GENDER==2 ~ "Female")) %>% 
  mutate(edu = case_when(
    DEGREE==0 | DEGREE==1 | DEGREE==2 ~ "HS or <",
    DEGREE==3 | DEGREE==4 | DEGREE==9 ~ "College/Some",
    DEGREE==5 | DEGREE==6 ~ "> College")) 
   
 
myvars_2 <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "stroke", "hbp", "t2db", "bmi", "smoking", "alcohol", "proxy_f", "totalscore", "baseline_crp", "baseline_hdl", "baseline_tc", "baseline_cysc", "baseline_a1c", "demdx", "APOE2010_bin") 
dem_data_ffff <- dplyr::select(dem_final_fff,myvars_2)  
```



# Sample Size 

```{r}

complete_data <- dem_data_ffff

complete_crp <- complete_data %>% 
  filter(baseline_crp!="NA")  # N = 12,394, we lost 670 from the 13,064
 
complete_hdl <- complete_crp %>% 
  filter(baseline_hdl!="NA") # N = 10,458 

complete_tc <- complete_hdl %>% 
   filter(baseline_tc!="NA") # N= 10,455

complete_cysc <- complete_tc %>% 
   filter(baseline_cysc!="NA") # N= 10,403

complete_a1c <- complete_cysc %>% 
  filter(baseline_a1c!="NA") # N=10,380

complete_data_biomarkers <- complete_data %>% 
filter(baseline_crp!="NA" & baseline_hdl!="NA" & baseline_tc!="NA" & baseline_cysc!="NA" & baseline_a1c!="NA") # same as complete_a1c, therefore we lost 2,694 without complete biomarker data 


complete_demographics <- complete_data_biomarkers %>% 
  filter(race_ff!="NA" & gender_f!="NA" & education!="NA" & age!="NA")  # N = 10,362 , we passed from 10380 to 10362, we lost 18 people 

complete_health <- complete_demographics %>% 
  filter(alcohol!="NA" & stroke!="NA", hbp!="NA", t2db!="NA", bmi!="NA", smoking!="NA") # N = 10,147 we lost 215

complete_outcomes <- complete_health %>%  
  filter(demdx!="NA" & totalscore!="NA") # N = 10,147

complete_genetics <- complete_outcomes %>% 
  filter(APOE2010_bin!="NA") # 9,298


# Using na.omit and not including the mutually exclusive outcomes variable only demdx and totalscore 

myvars_3 <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "stroke", "hbp", "t2db", "bmi", "smoking", "alcohol", "proxy_f", "totalscore", "baseline_crp", "baseline_hdl", "baseline_tc", "baseline_cysc", "baseline_a1c", "demdx") 

complete_data_filter <- dplyr::select(complete_data,myvars_3)  

data_nm <- complete_data_filter%>% na.omit() # I get the same number of observation with full sample 10,147

# Sample Including Genetics APOE-information 

myvars_4 <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "stroke", "hbp", "t2db", "bmi", "smoking", "alcohol", "proxy_f", "totalscore", "baseline_crp", "baseline_hdl", "baseline_tc", "baseline_cysc", "baseline_a1c", "demdx", "APOE2010_bin")

complete_data_filter_apoe <- dplyr::select(complete_data,myvars_4)  

data_nm_apoe <- complete_data_filter_apoe%>% na.omit() # I get full sample 9,298

#save(data_nm, file = "full_data.RData")
#save(data_nm_apoe, file = "full_data_apoe.RData")


```


# Importing Study Data Complete Sample Size and Creating Outcome Variables 


```{r}
# Data cleaning 

rm(list = ls())
setwd("~/Documents/Dementia project /NIH grant/Data")
load("full_data.RData")
full_data = data_nm

full_data <- full_data %>% 
 mutate(demdx_f = case_when(
    demdx==0 ~ "NORMAL",
    demdx==2 ~ "DEMENTIA", 
    demdx==1 ~ "NA"
  )) %>% 
  mutate(cinddx= case_when(
    demdx==0 ~ "NORMAL",
    demdx==1 ~ "CIND", 
    demdx==2 ~ "NA"
  )) %>% 
  mutate(demdx_f = na_if(demdx_f, "NA")) %>% 
  mutate(cinddx = na_if(cinddx, "NA")) 

full_data <- full_data %>% 
  mutate_if(is.character,as.factor) %>% 
  mutate(cinddx_f = case_when(
    cinddx=="NORMAL" ~ 0, 
    cinddx=="CIND" ~ 1
  )) %>% 
  mutate(demdx_ff = case_when(
    demdx_f=="NORMAL" ~ 0, 
    demdx_f=="DEMENTIA" ~ 1
  )) %>%  
  mutate(race_ff = na_if(race_ff, "Other-NH")) %>% 
   filter(race_ff!='NA')

```


# Quartiles of CysC Distribution in Study Sample 

```{r}


full_data %>% 
  summarize("Obs" = n(),
  "Mean" =mean(baseline_cysc,na.rm=TRUE),
  "SD" = sd(baseline_cysc,na.rm=TRUE),
  "Min" = min(baseline_cysc, na.rm=TRUE),
  "Max" = max(baseline_cysc, na.rm=TRUE),
  "25%" = quantile(baseline_cysc, probs=0.25, na.rm=TRUE), "50%" = quantile(baseline_cysc, probs=0.50, na.rm=TRUE), "75%" = quantile(baseline_cysc, probs=0.75, na.rm=TRUE),
  "IQR" = IQR(baseline_cysc, na.rm=TRUE)) %>% 
  na.omit() 

full_data <-  dplyr::rename(full_data, race=race_ff, gender=gender_f) 

full_data <- full_data %>% 
  mutate(cysc_cat = case_when(
    baseline_cysc <= 0.86 ~ 0, 
    baseline_cysc >0.86 & baseline_cysc <= 1 ~ 1,
    baseline_cysc > 1 & baseline_cysc <= 1.24 ~ 2, 
    baseline_cysc > 1.24 ~ 3
  )) %>%  
  mutate_at(vars(cysc_cat), as.factor) %>% 
    mutate(cysc_cat_75 = case_when(
    baseline_cysc <= 1.24 ~ 0, 
    baseline_cysc > 1.24 ~ 1
  )) %>%  
  mutate_at(vars(cysc_cat_75), as.factor)  %>%  
  mutate(anydem_1 = case_when(
    demdx_ff==1 | cinddx_f==1 ~ 1, 
    TRUE ~ 0 )) %>% 
    mutate_at(vars(anydem_1), as.factor) %>% 
  mutate(dicho_race = case_when(
    race=="White-NH" ~ 0, 
    race=="Black-NH" ~ 1
  )) %>%  
  mutate(dicho_race_h = case_when(
   race=="White-NH" ~ 0, 
   race=="Hispanic" ~ 1
  )) %>% 
  mutate_at(vars(dicho_race), as.factor) %>% 
  mutate_at(vars(dicho_race_h), as.factor) %>% 
  mutate(int_cysc = case_when(
    dicho_race==0 & cysc_cat_75==0 ~ 0, 
    dicho_race==0 & cysc_cat_75==1 ~ 1,
    dicho_race==1 & cysc_cat_75==0 ~ 2,
    dicho_race==1 & cysc_cat_75==1 ~ 3
  )) %>% 
  mutate_at(vars(int_cysc), as.factor) %>%  
  mutate(int_cysc_h = case_when(
    dicho_race_h==0 & cysc_cat_75==0 ~ 0, 
    dicho_race_h==0 & cysc_cat_75==1 ~ 1,
    dicho_race_h==1 & cysc_cat_75==0 ~ 2,
    dicho_race_h==1 & cysc_cat_75==1 ~ 3
  )) %>% 
  mutate_at(vars(int_cysc_h), as.factor)

full_data <- full_data %>% 
    mutate(cysc_race = case_when(
      int_cysc==0 ~ "White & Low CysC",
      int_cysc==1 ~ "White & High CysC",
      int_cysc==2 ~ "Black & Low CysC",
      int_cysc==3 ~ "Black & High CysC"
    )) %>% 
  mutate_at(vars(cysc_race), as.factor) %>% 
    mutate(cysc_h = case_when(
      int_cysc_h==0 ~ "White & Low CysC",
      int_cysc_h==1 ~ "White & High CysC",
      int_cysc_h==2 ~ "Hispanic & Low CysC",
      int_cysc_h==3 ~ "Hispanic & High CysC"
    )) %>% 
    mutate_at(vars(cysc_h), as.factor) %>% 
  mutate(baseline_tc_hdl=baseline_tc/baseline_hdl)



#save(full_data, file = "Model_Data_CysC.RData") #/*HERE THE DATA WAS SAVED*/

```


# Prepare Data for Genetic Model 

```{r}
rm(list = ls())
setwd("~/Documents/Dementia project /NIH grant/Data")
load("full_data_apoe.RData")

full_data_apoe = data_nm_apoe

full_data_apoe <- full_data_apoe %>% 
 mutate(demdx_f = case_when(
    demdx==0 ~ "NORMAL",
    demdx==2 ~ "DEMENTIA", 
    demdx==1 ~ "NA"
  )) %>% 
  mutate(cinddx= case_when(
    demdx==0 ~ "NORMAL",
    demdx==1 ~ "CIND", 
    demdx==2 ~ "NA"
  )) %>% 
  mutate(demdx_f = na_if(demdx_f, "NA")) %>% 
  mutate(cinddx = na_if(cinddx, "NA")) 

full_data_apoe <- full_data_apoe %>% 
  mutate_if(is.character,as.factor) %>% 
  mutate(cinddx_f = case_when(
    cinddx=="NORMAL" ~ 0, 
    cinddx=="CIND" ~ 1
  )) %>% 
  mutate(demdx_ff = case_when(
    demdx_f=="NORMAL" ~ 0, 
    demdx_f=="DEMENTIA" ~ 1
  )) %>% 
  mutate(race_ff = na_if(race_ff, "Other-NH")) %>% 
   filter(race_ff!='NA')

```


# Quartiles of CysC Distribution for APOE Study Sample 

```{r}

full_data_apoe %>% 
  summarize("Obs" = n(),
  "Mean" =mean(baseline_cysc,na.rm=TRUE),
  "SD" = sd(baseline_cysc,na.rm=TRUE),
  "Min" = min(baseline_cysc, na.rm=TRUE),
  "Max" = max(baseline_cysc, na.rm=TRUE),
  "25%" = quantile(baseline_cysc, probs=0.25, na.rm=TRUE), "50%" = quantile(baseline_cysc, probs=0.50, na.rm=TRUE), "75%" = quantile(baseline_cysc, probs=0.75, na.rm=TRUE),
  "IQR" = IQR(baseline_cysc, na.rm=TRUE)) %>% 
  na.omit() 

full_data_apoe <-  dplyr::rename(full_data_apoe, race=race_ff, gender=gender_f) 

full_data_apoe <- full_data_apoe %>% 
  mutate(cysc_cat = case_when(
    baseline_cysc <= 0.84 ~ 0, 
    baseline_cysc >0.84 & baseline_cysc <= 1 ~ 1,
    baseline_cysc > 1 & baseline_cysc <= 1.24 ~ 2, 
    baseline_cysc > 1.24 ~ 3
  )) %>%  
  mutate_at(vars(cysc_cat), as.factor) %>% 
    mutate(cysc_cat_75 = case_when(
    baseline_cysc <= 1.24 ~ 0, 
    baseline_cysc > 1.24 ~ 1
  )) %>%  
  mutate_at(vars(cysc_cat_75), as.factor)  %>%  
  mutate(anydem_1 = case_when(
    demdx_ff==1 | cinddx_f==1 ~ 1, 
    TRUE ~ 0 )) %>% 
    mutate_at(vars(anydem_1), as.factor) %>% 
  mutate(dicho_race = case_when(
    race=="White-NH" ~ 0, 
    race=="Black-NH" ~ 1
  )) %>%  
  mutate(dicho_race_h = case_when(
   race=="White-NH" ~ 0, 
   race=="Hispanic" ~ 1
  )) %>% 
  mutate_at(vars(dicho_race), as.factor) %>% 
  mutate_at(vars(dicho_race_h), as.factor) %>% 
  mutate(int_cysc = case_when(
    dicho_race==0 & cysc_cat_75==0 ~ 0, 
    dicho_race==0 & cysc_cat_75==1 ~ 1,
    dicho_race==1 & cysc_cat_75==0 ~ 2,
    dicho_race==1 & cysc_cat_75==1 ~ 3
  )) %>% 
  mutate_at(vars(int_cysc), as.factor) %>%  
  mutate(int_cysc_h = case_when(
    dicho_race_h==0 & cysc_cat_75==0 ~ 0, 
    dicho_race_h==0 & cysc_cat_75==1 ~ 1,
    dicho_race_h==1 & cysc_cat_75==0 ~ 2,
    dicho_race_h==1 & cysc_cat_75==1 ~ 3
  )) %>% 
  mutate_at(vars(int_cysc_h), as.factor)

full_data_apoe <- full_data_apoe %>% 
    mutate(cysc_race = case_when(
      int_cysc==0 ~ "White & Low CysC",
      int_cysc==1 ~ "White & High CysC",
      int_cysc==2 ~ "Black & Low CysC",
      int_cysc==3 ~ "Black & High CysC"
    )) %>% 
  mutate_at(vars(cysc_race), as.factor) %>% 
    mutate(cysc_h = case_when(
      int_cysc_h==0 ~ "White & Low CysC",
      int_cysc_h==1 ~ "White & High CysC",
      int_cysc_h==2 ~ "Hispanic & Low CysC",
      int_cysc_h==3 ~ "Hispanic & High CysC"
    )) %>% 
    mutate_at(vars(cysc_h), as.factor) %>% 
  mutate(baseline_tc_hdl=baseline_tc/baseline_hdl)


#save(full_data_apoe, file = "Model_Data_CysC_APOE.RData") #/*HERE THE DATA WAS SAVED*/

# Load data from APOE to export to STATA and calculate models 

rm(list = ls())


setwd("~/Documents/Dementia project /NIH grant/Data")
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 

load("Model_Data_CysC_APOE.RData")

full_data_apoe <- full_data_apoe %>% 
  mutate(smoke = case_when(
    smoking== 0 ~ "Never Smoker",
    smoking==1 ~ "Former Smoker",
    smoking==2 ~ "Current Smoker"
  )) 

full_data_apoe <- full_data_apoe %>% 
  mutate(hbpf = case_when(
    hbp==0 ~ "No HPB", 
    hbp==1 ~ "Yes HBP"
  )) %>% 
  mutate_at(vars(hbpf), as.character()) 

# Export data to STATA for building Poisson Models with Robust Variance and Calculating RERI and CMAD

#library(foreign)
#write.dta(full_data_apoe, "Model_Data_CysC_APOE_ForSTATA.dta") 



```


# Prepare data for the expert model 

```{r}

rm(list = ls())

setwd("~/Documents/Dementia project /NIH grant/Data")
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 

load("Model_Data_CysC.RData")

load("hrs_democoggen_wide20200910.Rda") 

myvars <- c("HHID", "PN","expert_dem_06", "expert_dem_08")

expert_data <- dplyr::select(hrs0910,myvars)
expert_data %>% 
  dim()

merged_expert <- left_join(full_data, expert_data, by = c("HHID" = "HHID", "PN" = "PN")) 

merged_expert %>% 
  dim()

# Creating dementia expert variable for each wave 

merged_expert <- merged_expert %>% 
   mutate(expert_dem= case_when(
    wave==2006 ~ expert_dem_06,
    wave==2008 ~ expert_dem_08)) 

table(merged_expert$expert_dem)
table(merged_expert$demdx_ff)

#save(merged_expert, file = "Model_Data_CysC_Expert.RData") #/*HERE THE DATA WAS SAVED*/

# Export data to STATA for building Merged Models with Robust Variance and Calculating RERI and CMAD

#load("Model_Data_CysC_Expert.RData")

#library(foreign)
#write.dta(merged_expert, "Model_Data_CysC_Expert_ForSTATA.dta") 

```



# Build data set for excluded sample 


```{r}

# Extract all biomarker data 

rm(list = ls())
setwd("~/Documents/Dementia project /NIH grant/Data")
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 
paste0(source_dir, "hrs_democoggen_wide20200910.Rda")
load("hrs_democoggen_wide20200910.Rda") 

load("biomk06.RData")
load("biomk08.RData")
biomk06 <- biomk06
biomk08 <- biomk08

myvars_bio_2006 <- c("HHID", "PN", "KA1C_ADJ", "KHDL_ADJ", "KTC_ADJ", "KCYSC_ADJ", "KCRP_ADJ") 
myvars_bio_2008 <- c("HHID", "PN", "LA1C_ADJ", "LHDL_ADJ", "LTC_ADJ", "LCYSC_ADJ", "LCRP_ADJ") 

biomk06_f <- dplyr::select(biomk06,myvars_bio_2006)  
biomk08_f <- dplyr::select(biomk08,myvars_bio_2008)

# Renaming Variables for each data set biomarker data 
#2006 
biomk06_baseline <- dplyr::rename(biomk06_f, baseline_a1c=KA1C_ADJ , baseline_hdl = KHDL_ADJ,  baseline_tc = KTC_ADJ, baseline_cysc = KCYSC_ADJ, baseline_crp = KCRP_ADJ)

biomk06_baseline <- biomk06_baseline %>% 
                    mutate(wave=2006)

# 2008 
biomk08_baseline <- dplyr::rename(biomk08_f, baseline_a1c = LA1C_ADJ, baseline_hdl = LHDL_ADJ, baseline_tc = LTC_ADJ, baseline_cysc = LCYSC_ADJ, baseline_crp = LCRP_ADJ)

biomk08_baseline <- biomk08_baseline %>% 
                    mutate(wave=2008)

biomarker_final <- rbind(biomk06_baseline, biomk08_baseline) 

# This data set should have 13,064 observation which is the original sample size of both DBS sample: the 2006 and 2008 combined into 1 file

biomarker_final %>% 
  dim()

# Merge the biomarker final data with the HRS repository created by Erin

merged_data_missing <- left_join(biomarker_final, hrs0910, by = c("HHID" = "HHID", "PN" = "PN")) 

# Creating the Outcome Variable

# Assigning Waves & Creating Outcome Variables 
merged_data_missing <- merged_data_missing %>% 
  mutate(AD_06 = case_when(
    AD06==3 ~ 0, 
    AD06==2 ~ 1,
    AD06==1 ~ 2
  )) %>% 
  mutate(AD_08 = case_when(
    AD08==3 ~ 0,
    AD08==2 ~ 1, 
    AD08==1 ~ 2
  )) %>% 
  mutate(demdx= case_when(
    wave==2006 ~ AD_06,
    wave==2008 ~ AD_08)) %>% 
  mutate(age = wave-BIRTHYR) %>% 
  mutate(education = SCHLYRS) %>% 
  mutate(stroke = case_when(
    wave==2006 ~ STROKE06,
    wave==2008 ~ STROKE08
  )) %>% 
  mutate(hbp = case_when(
    wave==2006 ~ R8HIBPE,
    wave==2008 ~ R9HIBPE # had a mistake in here as R8HIBPE, just corrected it 
  )) %>% 
  mutate(t2db = case_when(
    wave==2006 ~ R8DIABE,
    wave==2008 ~ R9DIABE
  )) %>% 
  mutate(bmi = case_when(
    wave==2006 ~ R8BMI,
    wave==2008 ~ R9BMI
  )) %>% 
  mutate(smoking = case_when(
    wave==2006 ~ smoke06,
    wave==2008 ~ smoke08
  )) %>% 
  mutate(alcohol = case_when(
    wave==2006 ~ R8DRINKN,
    wave==2008 ~ R9DRINKN
  )) %>% 
  mutate(proxy = case_when(
    wave==2006 ~ KA009,
    wave==2008 ~ LA009,
  )) %>% 
  mutate(proxy_f= case_when(
    proxy==1 ~ 1,
    proxy==2 ~ 2, 
    proxy==3 ~ 2
  )) %>% 
  mutate(totalscore= case_when(
    wave==2006 ~ totscore06,
    wave==2008 ~ totscore08
  )) %>% 
  mutate(race_f = case_when(
    RACE==0 ~ "NA",
    RACE==1 ~ "White",
    RACE==2 ~ "Black",
    RACE==7 ~ "Other")) %>% 
  mutate(hispanic_f = case_when(
      HISPANIC==1 | HISPANIC==2 | HISPANIC==3 ~ "Hispanic",
      HISPANIC==5 ~ "Non-Hispanic",
      HISPANIC==0 ~ "NA")) 


merged_data_missing <- merged_data_missing  %>% 
  mutate(race_f = na_if(race_f, "NA")) %>% 
  mutate(hispanic_f = na_if(hispanic_f, "NA")) %>% 
  mutate(race_ff = case_when(
    race_f=="White" & hispanic_f=="Non-Hispanic" ~ "White-NH",
    race_f=="Black" & hispanic_f=="Non-Hispanic" ~ "Black-NH",
    race_f=="Other" & hispanic_f=="Non-Hispanic" ~ "Other-NH",
    hispanic_f=="Hispanic" ~ "Hispanic")) %>% 
  mutate(gender_f = case_when(
    GENDER==1 ~ "Male",
    GENDER==2 ~ "Female")) %>% 
  mutate(edu = case_when(
    DEGREE==0 | DEGREE==1 | DEGREE==2 ~ "HS or <",
    DEGREE==3 | DEGREE==4 | DEGREE==9 ~ "College/Some",
    DEGREE==5 | DEGREE==6 ~ "> College")) %>% 
    mutate(race_ff = na_if(race_ff, "Other-NH")) %>% 
   filter(race_ff!='NA')
 
myvars_missing <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "stroke", "hbp", "t2db", "bmi", "smoking", "alcohol", "totalscore", "baseline_crp", "baseline_hdl", "baseline_tc", "baseline_cysc", "baseline_a1c", "demdx") 


merged_data_missing <- dplyr::select(merged_data_missing, myvars_missing)  

merged_data_missing_nm <- merged_data_missing %>% na.omit()

merged_data_missing_nm <- merged_data_missing_nm %>% 
                          mutate(missing=0)

myvars_missing_f <- c("HHID", "PN", "missing") 

merged_data_missing_nm <- dplyr::select(merged_data_missing_nm, myvars_missing_f) 

complete_data_mising <- left_join(merged_data_missing, merged_data_missing_nm, by = c("HHID" = "HHID", "PN" = "PN"))

complete_data_mising <- complete_data_mising %>% 
  mutate(missing_f = ifelse(is.na(missing), 1, 0))

```


# Missing Data Table 


```{r}


complete_data_mising <- complete_data_mising %>% 
  mutate(smoke = case_when(
    smoking== 0 ~ "Never Smoker",
    smoking==1 ~ "Former Smoker",
    smoking==2 ~ "Current Smoker"
  )) 

complete_data_mising <- complete_data_mising %>% 
  mutate(hbpf = case_when(
    hbp==0 ~ "No HPB", 
    hbp==1 ~ "Yes HBP"
  )) %>% 
  mutate_at(vars(hbpf), as.character()) %>% 
  mutate(baseline_tc_hdl=baseline_tc/baseline_hdl) %>% 
    mutate(demdx_f = case_when(
    demdx== 0 ~ "Normal",
    demdx==1 ~ "CIND",
    demdx==2 ~ "Dementia"
  )) 


myvars_missing <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "stroke", "hbp", "t2db", "bmi", "smoking", "alcohol", "totalscore", "baseline_crp", "baseline_hdl", "baseline_cysc", "baseline_a1c", "demdx") 



table1_data <- complete_data_mising %>%  
  dplyr::select("demdx_f", "race_ff", "gender_f", "edu","education",  "age","alcohol", "smoke", "bmi", "hbpf", "baseline_cysc","baseline_a1c", "baseline_tc_hdl" , "wave" , "missing_f")

table1_missing <- table1_data %>%
  tbl_summary(
    by = missing_f, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
     label = list(demdx_f ~ "Cognition",
                  race_ff ~ "Race",
                  gender_f ~ "Gender", 
                  edu ~ "Education Category",
                  education ~ "Years of Education",
                  age ~ "Age",
                  alcohol ~ "Alcohol",
                  smoke ~ "Smoking",
                  bmi ~ "Body Mass Index",
                  hbpf ~ "Hypertension" , 
                  baseline_cysc ~ "Baseline Cystatin C",
                 baseline_a1c ~ "Baseline HbA1C", 
                 baseline_tc_hdl ~ "Baseline TC/HDL-C", 
                 wave ~ "Wave"
  )) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table1_missing %>% as_flex_table



``` 




