---
title: Tables for Dementia, CIND, and Cystatin-C  
author: "César Higgins"
date: '`r Sys.Date()`'
output:   
  html_document:
    code_folding: hide
    highlight: pygments
    keep_md: no
    number_sections: yes
    smart: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(compareGroups)
library(knitr)
library(kableExtra)
library(janitor)
library(broom)
library(lubridate)
library(magrittr) # need to run every time you start R and want to use %>%
library(kableExtra)
library(gtsummary)
library(foreign)
library(gmodels)
library(stringr)
library(dplyr)  
library(flextable) # needs this package for creating flexible tables for gtsummary tables and export to word 

summarize = dplyr::summarize

knitr::opts_chunk$set(
  eval       = TRUE,    # whether to run code in code chunk
  include    = TRUE,    # whether to include the chunk output
  echo       = TRUE,   # Whether to show code chunk in final output
  error      = TRUE,    # whether to display error messages
  message    = FALSE,   # whether to preserve messages
  warning    = FALSE,   # whether to preserve warnings
  comment    = "#>",    # a character string to append at start
                        # of each line of results in final document
  tidy       = FALSE,   # whether to tidy code chunks for display
  #dpi        = 96, 
  #fig.width  = 6,       # consistent width for figures
  #fig.asp    = 0.618,   # the golden ratio, can be adjusted in individual chunks
  out.width  = "100%",   # controls the output size
  fig.align  = "center" # give plot room to breathe
)

```


```{r read data, results="hide"}
rm(list = ls())
setwd("~/Documents/Dementia project /NIH grant/Data")
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 
paste0(source_dir, "Model_Data_CysC.RData")
load("Model_Data_CysC.RData")

```



# Tables {.tabset}

## Table 1

Sample characteristics by the percentiles of cystatin C distribution, United States Health and Retirement Study, 2006 and 2008

```{r Table 1, eval=TRUE}

# Smoking and hypertension make them factors for this table

full_data <- full_data %>% 
  mutate(smoke = case_when(
    smoking== 0 ~ "Never Smoker",
    smoking==1 ~ "Former Smoker",
    smoking==2 ~ "Current Smoker"
  )) 

 full_data <- full_data %>% 
  mutate(hbpf = case_when(
    hbp==0 ~ "No HPB", 
    hbp==1 ~ "Yes HBP"
  )) %>% 
   mutate(race_f = case_when(
     race=="Black-NH" ~ "Black-NH",
     race=="Hispanic" ~ "Hispanic", 
     race=="White-NH" ~ "White-NH"
   ))

# Export data to STATA for building Poisson Models with Robust Variance and Calculating RERI and CMAD

#save(full_data, file = "Model_Data_CysC_ForSTATA.RData")
#load("Model_Data_CysC_ForSTATA.RData")
#library(foreign)
#write.dta(full_data, "Model_Data_CysC_ForSTATA.dta") 

table1_data <- full_data %>%  
  dplyr::select("cysc_cat", "demdx_f", "cinddx", "race_f", "gender", "edu", "age", "education", "alcohol", "smoke", "bmi", "hbpf", "baseline_a1c", "baseline_tc_hdl")



table1_cysc <- table1_data %>%
  tbl_summary(
    by = cysc_cat, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
     label = list(demdx_f ~ "Dementia",
                cinddx ~ "CIND", 
                gender ~ "Gender", 
                race_f ~ "Race",
                edu ~ "Education", 
                education ~ "Years of Education", 
                alcohol ~ "Alcohol", 
                smoke ~ "Smoking Status", 
                bmi ~ "Body Mass Index", 
                age ~ "Age",
                baseline_a1c ~ "Baseline A1C", 
                baseline_tc_hdl ~ "Baseline TC/HDL-C",
                hbpf ~ "High Blood"
  )) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table1_cysc_f <- table1_cysc %>%  as_flex_table()

table1_cysc_f

```


# Supplemental Tables {.tabset}


## Supplemental Table 2 

Distribution of sample characteristics by the joint effect of race/ethnicity (non-Hispanic Black vs non-Hispanic White) and high cystatin C levels (>1.24mg/L) in the United States Health and Retirement Study, waves 2006 and 2008

```{r Table for blacks and CysC, eval=TRUE}

table1a_data <- full_data %>%  
  dplyr::select("cysc_race", "demdx_f", "cinddx", "gender", "edu", "age", "education", "alcohol", "smoke", "bmi", "hbpf", "baseline_a1c", "baseline_tc_hdl") 

table1a_cysc <- table1a_data %>% 
tbl_summary(
    by = cysc_race, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
   label = list(demdx_f ~ "Dementia",
                cinddx ~ "CIND", 
                gender ~ "Gender", 
                edu ~ "Education", 
                education ~ "Years of Education", 
                alcohol ~ "Alcohol", 
                smoke ~ "Smoking Status", 
                bmi ~ "Body Mass Index", 
                age ~ "Age",
                hbpf ~ "Hypertension",
                baseline_a1c ~ "Baseline A1C", 
                baseline_tc_hdl ~ "Baseline TC/HDL-C"
  )) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table1a_cysc %>% as_flex_table

```


## Supplemental Table 3 

Distribution of sample characteristics by the joint effect of ethnicity (Hispanic vs non-Hispanic White) and high cystatin C levels (>1.24mg/L) in the United States Health and Retirement Study, waves 2006 and 2008

```{r Table for Hispanics and CysC, eval=TRUE}

table1b_data <- full_data %>%  
  dplyr::select("cysc_h", "demdx_f", "cinddx","gender", "edu", "age", "education", "alcohol", "smoke", "bmi", "hbpf", "baseline_a1c", "baseline_tc_hdl")

table1b_cysc <- table1b_data %>%
  tbl_summary(
    by = cysc_h, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
   label = list(demdx_f ~ "Dementia",
                cinddx ~ "CIND", 
                gender ~ "Gender", 
                edu ~ "Education", 
                education ~ "Years of Education", 
                alcohol ~ "Alcohol", 
                smoke ~ "Smoking Status", 
                bmi ~ "Body Mass Index", 
                age ~ "Age",
                hbpf ~ "Hypertension",
                baseline_a1c ~ "Baseline A1C", 
                baseline_tc_hdl ~ "Baseline TC/HDL-C"
  )) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table1b_cysc %>% as_flex_table



```



# Figures {.tabset}

## Density Plot 

Distribution of natural logarithm transformation of cystatin C using a Kernel density plot. The 75th percentile (cystatin C serum levels > 1.24mg/L) is indicated by the dotted line. 

```{r, eval=TRUE}

full_data <- full_data %>%  
  mutate(dem_plot = case_when(
    demdx_ff==0 ~ 0,
    demdx_ff==1 ~ 1
  )) 


summary_data <-  full_data %>% 
  summarize(top_p = log(quantile(baseline_cysc, probs=0.75, na.rm=TRUE)))%>% 
  na.omit() 

# Density Plot Figure

cysc <- ggplot(full_data, aes(x=log(baseline_cysc))) +
  geom_density(alpha=0.8, fill="lightblue") + 
geom_vline(data=summary_data, aes(xintercept=top_p), linetype="dashed") +
  labs(title="Baseline Cystatin C",x="Natural Log of Cystatin C (mg/L)", y = "Density") +
  theme_classic()

cysc


```

## Plot of predicted probabilities 


Marginal predicted probabilities of high cystatin C (>1.24mg/L) stratified by race/ethnicity in different age groups, in the Health and Retirement Study waves 2006 & 2008.

```{r}

library(readxl)
probabilities <- read_excel("~/Documents/Dementia project /NIH grant/Data/probabilities.xlsx")
View(probabilities)  

legend_title <- "Race/Ethnicity"

p_7 <- ggplot(probabilities, aes(x=age_group, y=predicted_p, group=Race, color=Race))+ 
    geom_errorbar(aes(ymin=l_ci, ymax=u_ci), width=.05, 
    position=position_dodge(0.03)) +
    geom_line(aes(linetype=Race)) + 
    geom_point(aes(shape=Race))+
    labs(title="Predictive Margins with 95% CIs",x= "Age Groups", y ="Predicted Probability of High Cystatin C")+
  scale_color_brewer(palette="Paired") +
    theme(text = element_text(size = 16))

p_7 

```